Stack Safety as a Security Property

### 2020-07-01 main.tex
Comments / polishing

\begin{figure}
  \begin{minipage}{\textwidth}
    \begin{center}
\MemoryLabel{10em}{1.5em}{40}
\MemoryLabel{16em}{1.5em}{60}
\MemoryLabel{22em}{1.5em}{80}
$\cdots$
\memory{4}{gray}%
\memory{5}{green}[{\makebox[0pt]{Caller}}]%
\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
\memory{8}{red}[{\makebox[0pt]{Callee}}]%
\memory{4}{gray}%
~$\cdots$\\
  \end{center}
  \end{minipage}\\
~\\
~\\
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
lw r1 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}~
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
sw r0 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}
  \caption{Examples of Unsafe Programs}
  \label{fig:stackunsafety}
\end{figure}

@@

The call stack is an increasingly popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...} aiming to achieve {\em stack
  safety}.

But what {\em is} stack safety? A precise formal definition of stack
safety is lacking from the literature, with the notion only being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other approaches}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new part?}
Recently, a \leo{inspiring/awesome/something good} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
while impressive, stack safety was defined with respect to an abstract
machine that ``exhibits stack safety by construction''. Which begs
the question, what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. In this paper
we turn this intuition into two formal \leo{extensional} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully developed.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property, with a confidentiality and an
integrity component.
%
At a high level, {\em stack confidentiality} ensures that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that if a function call returns, the caller's
stack data will be identical to what it was before the call.

While integrity and confidentiality are relatively standard security
properties, this setting provides an interesting additional challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be potentially proteced from {\em its} callees.  In
this paper we address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms.

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code either
directly (via a jump instruction) or indirectly (by corrupting stack
data and returning). \leo{There needs to be some informal description
  of the wbcf part here. Sean?}

We begin by giving some background in Section~\ref{sec:prelim}.
Our contributions are:
\begin{itemize}
\item \leo{More? Extensional} formal characterizations of the
  intuitive concepts of local state encapsulation~\ref{sec:lse} and
  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
  property is strong enough to capture the folk intuition: we show
  that the stack unsafe examples from the literature also fail our
  property.
\item Evidence that the property is not too strong. It's realizible by
  a previously proposed micropolicy enforcement
  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
    substantial body of code} We then present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker.
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework.
\end{itemize}
Finally, in Section~\ref{sec:relwork} we discuss related work and
sketch directions for future work in Section~\ref{sec:future}.

>>>>

\begin{figure}
  \begin{minipage}{\textwidth}
    \begin{center}
\MemoryLabel{10em}{1.5em}{40}
\MemoryLabel{16em}{1.5em}{60}
\MemoryLabel{22em}{1.5em}{80}
$\cdots$
\memory{4}{gray}%
\memory{5}{green}[{\makebox[0pt]{Caller}}]%
\memory{5}{yellow}[{\makebox[0pt]{Args}}]%
\memory{8}{red}[{\makebox[0pt]{Callee}}]%
\memory{4}{gray}%
~$\cdots$\\
  \end{center}
  \end{minipage}\\
~\\
~\\
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
lw r1 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}~
  \begin{minipage}{0.45\textwidth}
    \begin{center}
\begin{verbatim}
jal ...
sw r0 r0 40
\end{verbatim}
    \end{center}
  \end{minipage}
  \caption{Examples of Unsafe Programs\bcp{The numbers don't line up with
      the boxes; the program fragments need comments explaining what each
      line does.}}
  \label{fig:stackunsafety}
\end{figure}

@@

The call stack is an increasingly popular target for attacks
exploiting memory safety vulnerabilities, with potential consequences
ranging from attackers reading or corrupting private stack data, to
highjacking a program's control flow. To prevent such attacks, many
techniques have been proposed in the literature\leo{cite, point out?
  stack canaries, cherry, nicks...}\bcp{I think pointing out makes a
  stronger point.} aiming to achieve ``stack
  safety''.

But what {\em is} this stack safety of which we speak? A precise formal
definition is lacking in the literature, with the notion usually being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other
  approaches}\bcp{Yes, we have to be very careful about this}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new
  part?}\bcp{If it is not too complicated to explain, we should definitely
  put it
  here.  If it is complicated, we can just say in text that there are other
  sorts of attacks too, and that we deal with them, and give a forward
  pointer.}
Recently, a \leo{inspiring/awesome/something good}\bcp{Maybe just say that
  we are only aware of one, quite recent, attempt to do something better} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
but even here, stack safety is defined with respect to an abstract
machine that ``exhibits stack safety by construction,'' begging
the question of what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. The first
contribution of the present paper is to
turn this intuition into two formal \leo{extensional}\bcp{Not sure we can
  claim this} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully
developed.}\bcp{Yes, and let's remember that they are highly likely to be
reviewing this paper!  Let's be as generous as we can be without
under-selling our delta.}

The underlying notion behind local state encapsulation is the
intention to protect callers from callees: a callee should be not be
able to access or overwrite a caller's private data. We formalize this
intuition as a security property with both confidentiality and
integrity components.
%
Intuitively, {\em stack confidentiality} promises that a callee's
observable behavior is not affected by memory outside its stack frame
(e.g. a caller's private data or uninitialized memory).  {\em Stack
  integrity} guarantees that, if a function call ever returns, the caller's
stack data at the point of return will be identical to what it was before
the call.  \bcp{We need to point out somewhere that we are only dealing with
stack safety, though we discuss (somewhere) how our property can be combined
with the heap safety property of [Azevedo et al]}

While integrity and confidentiality are relatively standard security
properties, the stack-safety setting raises an interesting additional
challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be proteced from {\em its} callees.
We address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms. \bcp{The two parts of
  the sentence are about two very different things.  The second deserves its
  own paragraph, IMO.}

Similarly, the underlying notion behind well-bracketed control flow is
the need to protect the call structure itself: a callee should not be
allowed to jump to the middle of an arbitrary block of code, either
directly (via a jump instruction) or indirectly (by corrupting the return
address on the stack and then returning). \leo{There needs to be some
  informal description of the wbcf part here. Sean?}

In summary, we offer the following contributions:
\begin{itemize}
\item \leo{More? Extensional} formal characterizations of the
  intuitive concepts of local state encapsulation~\ref{sec:lse} and
  well-bracketed control flow~\ref{sec:wbcf}.  Evidence that this
  property is strong enough to capture the folk intuition: we show
  that the stack unsafe examples from the literature also fail our
  property.
\item Evidence that the property is not too strong. It's realizible by
  a previously proposed micropolicy enforcement
  mechanism~\ref{sec:micropolicy}. \leo{And it says yes on a
    substantial body of code} We then present a variation of the
  property + policy that is lazy, easier/more efficient to implement,
  and provably weaker.
\item A provably stronger (inductive? step-by-step/stepwise) variant
  of this property tuned for testing/verification.  We test the whole
  thing with a cool framework.
\end{itemize}
We begin with some technical preliminaries in
Section~\ref{sec:prelim}. Section~\ref{sec:relwork} discusses related work, and
Section~\ref{sec:future} sketches directions for future work.
